Convex Surrogates via Polyhedral Losses

### 2019-02-02 cvx-fin.tex
cor 22

\begin{proposition}\label{prop:embed-trim}
  Let $\Gamma:\simplex\toto\reals^d$ be an elicitable property.
  The following are equivalent:
  \begin{enumerate}
  \item $\Gamma$ embeds a finite property $\gamma:\simplex \toto \R$.
  \item $\trim(\Gamma)$ is a finite set, and $\cup\,\trim(\Gamma) = \simplex$.
  \item There is a finite set of full-dimensional level sets $\Theta$ of $\Gamma$, and $\cup\,\Theta = \simplex$.
  \end{enumerate}
  Moreover, when any of the above hold, $\{\gamma_r : r\in\R\} = \trim(\Gamma) = \Theta$, and $\gamma$ is elicitable.
\end{proposition}
\begin{proof}
  Let $L$ elicit $\Gamma$.

@@

\begin{theorem}\label{thm:finite-elicit-embed}
  Let $\gamma$ be a finite property.
  The following are equivalent.
  \begin{enumerate}
  \item $\gamma$ is elicitable.
  \item $\gamma$ is embeddable.
  \item $\gamma$ is embeddable via a polyhedral loss.
  \end{enumerate}
\end{theorem}
\begin{proof}
  We trivially have $3\Rightarrow 2$, and by Proposition~\ref{prop:embed-trim} we also have $2\Rightarrow 1$.
  It thus suffices to show $1\Rightarrow 3$.
  In fact, this direction is a corollary of a stronger result given as Theorem~\ref{thm:general-duality-embedding} in the appendix, which states as follows.
  For any non-redundant property $\gamma'$, even if not finite, elicited by a loss with a Bayes risk which is a closed function, $\gamma'$ is embeddable in $d =n-1$ dimensions.
  By Lemma~\ref{lem:finite-full-dim}, the (negative) Bayes risk $G$ of the given finite elicitable property $\gamma$ is polyhedral and therefore closed, so the theorem applies.
  Moreover, the construction given in Theorem~\ref{thm:general-duality-embedding} defines the loss embedding $\gamma$ as $L(u) = C(u)\ones - u$ where $C=G^*$ is the convex conjugate of $G$, which is also polyhedral~\citet[Theorem 19.2]{rockafellar1997convex}.
  (Here $\ones\in\reals^\Y$ denotes the all-ones vector.)
  Thus $L$ is polyhedral.
\end{proof}

@@

\begin{theorem}\label{thm:polyhedral-embed}
  Every polyhedral loss embeds a finite elicitable property.
  Moreover, a polyhedral loss $L$ indirectly elicits a non-redundant elicitable property $\gamma$ if and only if $\gamma$ is finite and $L$ embeds a property which refines $\gamma$.
\end{theorem}
\begin{proof}
  Let $L:\reals^d\to\reals_+$ be a polyhedral loss.
  For all $p$, let $P(p)$ be the epigraph of the convex function $u\mapsto \inprod{p}{L(u)}$.
  From Lemma~\ref{lem:polyhedral-pd-same}, we have that the power diagram induced by the projection of $P(p)$ onto $\reals^d$ is constant whenever $p\in\inter\simplex$.
  Let $q\in\inter\simplex$ be the uniform distribution on $\Y$, and $V_\Y$ be the set of vertices of $P(q)$ projected onto $\reals^d$.
  By the above, this set is the same had we replaced $q$ by any $p\in\inter\simplex$.

@@

\begin{theorem}\label{thm:1d-tfae}
  Let $\gamma$ be a finite elicitable property.
  The following are equivalent.
  \begin{enumerate}
  \item $\gamma$ is orderable.
  \item $\elicembed(\gamma)=1$. ($\gamma$ is 1-embeddable.)
  \item $\elicpoly(\gamma)=1$. ($\gamma$ is indirectly elicitable via a polyhedral loss $L:\reals\to\reals^\Y$.)
  \item $\eliccvx(\gamma)=1$. ($\gamma$ is indirectly elicitable via a convex loss $L:\reals\to\reals^\Y$.)
  \item $\eliccts(\gamma)=1$. ($\gamma$ is indirectly elicitable via a real-valued continuous non-locally-constant property.)
  \end{enumerate}
\end{theorem}

@@

From Theorem~\ref{thm:1d-tfae}, we can already prove lower bounds on embedding dimension, as well as the other complexity classes mentioned.
Here are two simple examples.
\begin{corollary}
  \raf{Let $\gamma$ be abstain for any $\alpha$, or $\mathrm{mode}$, and any $n\geq 3$.  Then
    $\elicembed(\gamma)\geq 2$,
    $\elicpoly(\gamma)\geq 2$,
    $\eliccvx(\gamma)\geq 2$,
    $\eliccts(\gamma)\geq 2$.}
\end{corollary}
\begin{proof}
  % Every cell (level set) of an orderable property is adjacent to at most two other cells.
  % This fails to hold for \raf{abstain} when $n\geq 3$ and the mode when $n\geq 4$.
  % For the mode when $n=3$, we simply observe that the intersection of any two adjacent cells fails to be a hyperplane.
  The cell adjacency graph of an orderable property must be a path, but for the mode it is the complete graph on $n$ nodes, and for \raf{abstain} a star graph on $n+1$ nodes, which are not paths when $n \geq 3$.
\end{proof}

@@

Orderability is a useful characterization of 1-embeddable properties, in the sense that the orderability of a property is easily tested.
We seek a similar condition for higher dimensions.
Toward this end, let us decompose orderability into two conditions which will lend themselves to generalization.
A property $\gamma$ is orderable if and only if there exist closed intervals $[a(r)_y,b(r)_y] \subseteq \reals$ for all $r\in\R, y\in\Y$, such that the following two conditions hold:
\begin{enumerate}
\item (Optimality) For all $r\in\R, p\in\simplex$, we have $r\in \gamma(p) \iff \inprod{a(r)}{p} \leq 0 \leq \inprod{b(r)}{p}$.
\item (Monotonicity) There exists an ordering $\R=\{r_1,\ldots,r_k\}$ such that for all $y\in\Y$, $i\in\{1,\ldots,k-1\}$, we have $b(r_i) = a(r_{i+1})$.
\end{enumerate}
Thinking of the intervals $[a(r)_y,b(r)_y]$ as subgradients of a convex function $L_y$, the first condition states that the report $r$ is optimal exactly when it should be, and the second states that the intervals are indeed subgradients of some convex function.

Generalizing the above to higher dimensions, we obtain the following.
\begin{theorem} \label{thm:char}
  A finite property $\gamma:\simplex\toto\R$ is $d$-embeddable if and only if there exist polytopes $T(r,y) \subseteq \reals^d$ for all $r\in\R, y\in\Y$, such that the following two conditions hold:
  \begin{enumerate}
  \item (Optimality) For all $r\in\R, p\in\simplex$, we have $r\in \gamma(p) \iff 0 \in \sum_{y\in\Y} p_y T(r,y)$, where summation is the Minkowski sum.
  \item (Monotonicity) There exists an embedding $\varphi: \R \to \reals^d$ and set of convex functions $\{L_y:\reals^d\to\reals\}_{y\in\Y}$ such that for all $y\in\Y, r\in\R$ we have $T(r,y) = \partial L_y(\varphi(r))$.
  \end{enumerate}
\end{theorem}

@@

\bo{A whole section of the following proof can be cut out, but it's trying to be a tutorial. Pairs well with the figure (do we have a figure?)}
\begin{theorem}\label{thm:abs-5outcome-not-2embed}
  Abstain with $\alpha=1/2$ and $|\Y|=5$ is not $2$-embeddable.
\end{theorem}
\begin{proof}
By Corollary~\ref{cor:d-embed-implies-d-construction}, we can show there is no $2$-construction for one of the level sets.
We choose the abstain cell, which is $\{p : \max_y p_y \leq 0.5\}$.
Because this equals $\{p : (\forall y) 2p_y \geq \sum_{y'} p_{y'}$\}, we can write the level set as $\{p : Bp \geq 0\}$ for
  \[ B = \begin{pmatrix}
       -1 &  1 &  1 & \cdots & 1  \\
       1  & -1 &  1 & \cdots & 1  \\
          &    &    & \vdots &    \\
       1  &  1 &  1 & \cdots & -1
     \end{pmatrix} .
  \]
In other words, $B := \ones - 2I$, where $\ones$ is the all-ones matrix and $I$ is the identity.
In our example, $|\Y| = 5$, so $B$ has width $5$ because there are $5$ outcomes and height $5$ because there are $5$ constraints (one per outcome).

@@

\begin{theorem}
  Abstain with $\alpha > 1/2$ and $|\Y|=7$ is not $2$-embeddable.
\end{theorem}
\begin{proof}
  Like Theorem~\ref{thm:abs-5outcome-not-2embed}, we will again show that there is no $2$-construction of the abstain cell.
  We then write the matrix $B$ describing the abstain cell as follows:
	\[ B = \begin{pmatrix}
       -\beta &  1 &  1 & \cdots & 1  \\
       1  & -\beta &  1 & \cdots & 1  \\
          &    &    & \vdots &    \\
       1  &  1 &  1 & \cdots & -\beta
     \end{pmatrix} .
  \]

@@

@@

>>>>

\begin{proposition}\label{prop:embed-trim}
  Let $\Gamma:\simplex\toto\reals^d$ be an elicitable property.
  The following are equivalent:
  \begin{enumerate}\setlength{\itemsep}{0pt}
  \item $\Gamma$ embeds a finite property $\gamma:\simplex \toto \R$.
  \item $\trim(\Gamma)$ is a finite set, and $\cup\,\trim(\Gamma) = \simplex$.
  \item There is a finite set of full-dimensional level sets $\Theta$ of $\Gamma$, and $\cup\,\Theta = \simplex$.
  \end{enumerate}
  Moreover, when any of the above hold, $\{\gamma_r : r\in\R\} = \trim(\Gamma) = \Theta$, and $\gamma$ is elicitable.
\end{proposition}
\begin{proof}
  Let $L$ elicit $\Gamma$.

@@

\begin{theorem}\label{thm:finite-elicit-embed}
  Let $\gamma$ be a finite property.
  The following are equivalent.
  \begin{enumerate}\setlength{\itemsep}{0pt}
  \item $\gamma$ is elicitable.
  \item $\gamma$ is embeddable.
  \item $\gamma$ is embeddable via a polyhedral loss.
  \end{enumerate}
\end{theorem}
\begin{proof}
  We trivially have $3\Rightarrow 2$, and by Proposition~\ref{prop:embed-trim} we also have $2\Rightarrow 1$.
  It thus suffices to show $1\Rightarrow 3$.
  In fact, this direction is a corollary of a stronger result given as Theorem~\ref{thm:general-duality-embedding} in the appendix, which states as follows.
  For any non-redundant property $\gamma'$, even if not finite, elicited by a loss with a Bayes risk which is a closed function, $\gamma'$ is embeddable in $d =n-1$ dimensions.
  By Lemma~\ref{lem:finite-full-dim}, the (negative) Bayes risk $G$ of the given finite elicitable property $\gamma$ is polyhedral and therefore closed, so the theorem applies.
  Moreover, the construction given in Theorem~\ref{thm:general-duality-embedding} defines the loss embedding $\gamma$ as $L(u) = C(u)\ones - u$ where $C=G^*$ is the convex conjugate of $G$, which is also polyhedral~\citet[Theorem 19.2]{rockafellar1997convex}.
  (Here $\ones\in\reals^\Y$ denotes the all-ones vector.)
  Thus $L$ is polyhedral.
\end{proof}

@@

\begin{theorem}\label{thm:polyhedral-embed}
  Every polyhedral loss embeds a finite elicitable property.
  Moreover, a polyhedral loss $L$ indirectly elicits a non-redundant elicitable property $\gamma$ if and only if $\gamma$ is finite and $L$ embeds a property which refines $\gamma$.
\end{theorem}

@@

\begin{theorem}\label{thm:1d-tfae}
  Let $\gamma$ be a finite elicitable property.
  The following are equivalent.
  \begin{enumerate}\setlength{\itemsep}{0pt}
  \item $\gamma$ is orderable.
  \item $\elicembed(\gamma)=1$. ($\gamma$ is 1-embeddable.)
  \item $\elicpoly(\gamma)=1$. ($\gamma$ is indirectly elicitable via a polyhedral loss $L:\reals\to\reals^\Y$.)
  \item $\eliccvx(\gamma)=1$. ($\gamma$ is indirectly elicitable via a convex loss $L:\reals\to\reals^\Y$.)
  \item $\eliccts(\gamma)=1$. ($\gamma$ is indirectly elicitable via a real-valued continuous non-locally-constant property.)
  \end{enumerate}
\end{theorem}

@@

From Theorem~\ref{thm:1d-tfae}, we can already prove lower bounds on embedding dimension, as well as the other complexity classes mentioned.
Here are two simple examples.
\begin{corollary}
  Let $n\geq 3$ and $\alpha\in(0,1)$.  Then the following are bounded below by 2:
    $\elicembed(\abstain{\alpha})$,
    $\elicpoly(\abstain{\alpha})$,
    $\eliccvx(\abstain{\alpha})$,
    $\eliccts(\abstain{\alpha})$,
    $\elicembed(\mode)$,
    $\elicpoly(\mode)$,
    $\eliccvx(\mode)$,
    $\eliccts(\mode)$.
\end{corollary}
\begin{proof}
  % Every cell (level set) of an orderable property is adjacent to at most two other cells.
  % This fails to hold for \raf{abstain} when $n\geq 3$ and the mode when $n\geq 4$.
  % For the mode when $n=3$, we simply observe that the intersection of any two adjacent cells fails to be a hyperplane.
  The cell adjacency graph of an orderable property must be a path, but for the mode it is the complete graph on $n$ nodes, and for $\abstain{\alpha}$ a star graph on $n+1$ nodes, which are not paths when $n \geq 3$.
\end{proof}

@@

Orderability is a useful characterization of 1-embeddable properties, in the sense that the orderability of a property is easily tested.
We seek a similar condition for higher dimensions.
Toward this end, let us decompose orderability into two conditions which will lend themselves to generalization.
A property $\gamma$ is orderable if and only if there exist closed intervals $[a(r)_y,b(r)_y] \subseteq \reals$ for all $r\in\R, y\in\Y$, such that the following two conditions hold:
\begin{enumerate}\setlength{\itemsep}{0pt}
\item (Optimality) For all $r\in\R, p\in\simplex$, we have $r\in \gamma(p) \iff \inprod{a(r)}{p} \leq 0 \leq \inprod{b(r)}{p}$.
\item (Monotonicity) There exists an ordering $\R=\{r_1,\ldots,r_k\}$ such that for all $y\in\Y$, $i\in\{1,\ldots,k-1\}$, we have $b(r_i) = a(r_{i+1})$.
\end{enumerate}
Thinking of the intervals $[a(r)_y,b(r)_y]$ as subgradients of a convex function $L_y$, the first condition states that the report $r$ is optimal exactly when it should be, and the second states that the intervals are indeed subgradients of some convex function.

Generalizing the above to higher dimensions, we obtain the following.
\begin{theorem} \label{thm:char}
  A finite property $\gamma:\simplex\toto\R$ is $d$-embeddable if and only if there exist polytopes $T(r,y) \subseteq \reals^d$ for all $r\in\R, y\in\Y$, such that the following two conditions hold:
  \begin{enumerate}\setlength{\itemsep}{0pt}
  \item (Optimality) For all $r\in\R, p\in\simplex$, we have $r\in \gamma(p) \iff 0 \in \sum_{y\in\Y} p_y T(r,y)$, where summation is the Minkowski sum.
  \item (Monotonicity) There exists an embedding $\varphi: \R \to \reals^d$ and set of convex functions $\{L_y:\reals^d\to\reals\}_{y\in\Y}$ such that for all $y\in\Y, r\in\R$ we have $T(r,y) = \partial L_y(\varphi(r))$.
  \end{enumerate}
\end{theorem}

@@

\bo{A whole section of the following proof can be cut out, but it's trying to be a tutorial. Pairs well with the figure}
\begin{theorem}\label{thm:abs-5outcome-not-2embed}
  Abstain with $\alpha=1/2$ and $|\Y|=5$ is not $2$-embeddable.
\end{theorem}

@@

\begin{theorem}\label{thm:abstain-alphaghalf-not-2e}
  Abstain with $\alpha > 1/2$ and $|\Y|=6$ is not $2$-embeddable.
\end{theorem}

@@

\begin{proof}[Proof of Theorem~\ref{thm:polyhedral-embed}]
  Let $L:\reals^d\to\reals_+$ be a polyhedral loss.
  For all $p$, let $P(p)$ be the epigraph of the convex function $u\mapsto \inprod{p}{L(u)}$.
  From Lemma~\ref{lem:polyhedral-pd-same}, we have that the power diagram induced by the projection of $P(p)$ onto $\reals^d$ is constant whenever $p\in\inter\simplex$.
  Let $q\in\inter\simplex$ be the uniform distribution on $\Y$, and $V_\Y$ be the set of vertices of $P(q)$ projected onto $\reals^d$.
  By the above, this set is the same had we replaced $q$ by any $p\in\inter\simplex$.

@@

\section{Omitted Proofs}
\begin{proof}[Proof of Theorem~\ref{thm:abs-5outcome-not-2embed}]
By Corollary~\ref{cor:d-embed-implies-d-construction}, we can show there is no $2$-construction for one of the level sets.
We choose the abstain cell, which is $\{p : \max_y p_y \leq 0.5\}$.
Because this equals $\{p : (\forall y) 2p_y \geq \sum_{y'} p_{y'}$\}, we can write the level set as $\{p : Bp \geq 0\}$ for
  \[ B = \begin{pmatrix}
       -1 &  1 &  1 & \cdots & 1  \\
       1  & -1 &  1 & \cdots & 1  \\
          &    &    & \vdots &    \\
       1  &  1 &  1 & \cdots & -1
     \end{pmatrix} .
  \]
In other words, $B := \ones - 2I$, where $\ones$ is the all-ones matrix and $I$ is the identity.
In our example, $|\Y| = 5$, so $B$ has width $5$ because there are $5$ outcomes and height $5$ because there are $5$ constraints (one per outcome).

\begin{proof}[Proof of Theorem~\ref{thm:abstain-alphaghalf-not-2e}]
  Like Theorem~\ref{thm:abs-5outcome-not-2embed}, we will again show that there is no $2$-construction of the abstain cell.
  We then write the matrix $B$ describing the abstain cell as follows:
	\[ B = \begin{pmatrix}
       -\beta &  1 &  1 & \cdots & 1  \\
       1  & -\beta &  1 & \cdots & 1  \\
          &    &    & \vdots &    \\
       1  &  1 &  1 & \cdots & -\beta
     \end{pmatrix} .
  \]
